---
title: "Handout 4"
author: "Guillermo Gavilla Hernández, Miquel Toni Llambías Cabot"
output:
html_document
theme: united
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=8, fig.align="center", echo=TRUE, warning=FALSE, message=FALSE)
library(knitr)
library(tidyverse)
library(igraph)
```


## 1) Generate an undirected random graph using the model of "preferential attachment" (sample_pa( )) of 1000 nodes. With $\beta=0.1$, $\gamma=0.1$, generate a SIR pandemic (iterative method). The initial infected nodes should be the 10 highest using the eigen_centrality( ). Compare the results to when the initial nodes are at random. Reduce or increase $\beta$ and compare.

### Creamos las funciones

```{r sir_functions}
states <- c("susceptible", "infected", "recovered", "unknown")
add_states <- function (graph, susceptible, infected, rho) {
  if (!is.null(susceptible)) {
    V(graph)[susceptible]$state <- states[1]
    V(graph)[susceptible]$t_recover <- 0
  }
  if (!is.null(infected)) {
    V(graph)[infected]$state <- states[2]
    V(graph)[infected]$t_recover <- 1/rho
  }
  V(graph)[is.na(V(graph)$state)]$state <- states[4]
  return(graph)
}

do_if_infected <- function (graph, idx, beta, rho, tick) {
  vertex_neighbors <- neighbors(graph = graph, idx)
  for(i in as.numeric(vertex_neighbors)) {
    if (V(graph)$state[i] == states[1]) {
      random <- runif(1, 0, 1)[1]
      if (beta >= random) {
        V(graph)$state[i] <- states[2]
        V(graph)$t_recover[i] <- 1/rho + tick
      }
    }
  }
  return(graph)
}

change_state_infected <- function (graph, beta, rho, tick) {
  for(i in seq_along(V(graph))) {
    if (V(graph)$state[i] == states[2]) {
      graph <- do_if_infected(graph, i, beta, rho, tick)
      if (V(graph)$t_recover[i] <= tick) {
        V(graph)$state[i] <- states[3]
      }
    }
  }
  return(graph)
}

buildSummary <- function (graph, tick) {
  return(data.frame(
          time = tick,
          S = as.numeric(length(V(graph)[V(graph)$state == states[1]])),
          I = as.numeric(length(V(graph)[V(graph)$state == states[2]])),
          R = as.numeric(length(V(graph)[V(graph)$state == states[3]]))
  ))
}

plot_state <- function (graph, state) {
  V(graph)[V(graph)$state == states[1]]$color <- "SkyBlue2"
  V(graph)[V(graph)$state == states[2]]$color <- "red"
  V(graph)[V(graph)$state == states[3]]$color <- "green"
  plot.igraph(graph, vertex.color=V(graph)$color)
}

sir_2 <- function (graph, susceptible, infected, beta, rho, time, per_tick_plot = F) {
  graph <- add_states(graph, susceptible, infected, rho)
  per_tick_summary <- buildSummary(graph, 0)
  if (per_tick_plot){
    plot_state(graph, state)
  }
  for(tick in 1:time) {
    graph <- change_state_infected(graph, beta, rho, tick)
    per_tick_summary <- rbind(per_tick_summary, buildSummary(graph, tick))
    if (per_tick_plot){
      plot_state(graph, state)
    }
  }
  return(per_tick_summary)
}

print_sir_values_2 <- function (sirValues, title = "") {
  with(sirValues, {
    ggplot(data = sirValues, aes(x = time)) +
            ggtitle(title) +
            ylab("Nodes") +
            xlab("Time") +
            geom_line(aes(y = S), color = "blue") +
            geom_line(aes(y = I), color = "red") +
            geom_line(aes(y = R), color = "green") +
            scale_color_discrete(name = "State", labels = c("susceptible", "Infected", "Recovered"))
  })
}
```

#### Probamos que funciona

```{r sir_test}
sample_size <- 1000
infected_0 <- 10

prepare_data <- function (size, infected_0) {
  graph <- sample_pa(size, directed = F)
  eigen_values <- data.frame(id = 1: size, values = eigen_centrality(graph)$vector)
  infected <- ((eigen_values[order(eigen_values$values, decreasing = T),])[1:infected_0,])$id
  susceptible <- V(graph)[-infected]
  return(list(graph = graph, infected = infected, susceptible = susceptible))
}

set.seed(7288)
test_data <- prepare_data(30, 5)
sir_2(graph = test_data$graph,
                susceptible = test_data$susceptible,
                infected = test_data$infected,
                beta = 0.8,
                rho = 0.5,
                time = 5,
                per_tick_plot = T) %>%
  print_sir_values_2("test") %>%
  print()
```

#### Comparamos para diferentes betas

```{r sir_beta_compare}
graph_data <- prepare_data(size = sample_size, infected_0 = infected_0)
for (beta in c(0.01, 0.1, 0.3, 0.5))
{
  sir_2(graph = graph_data$graph,
        susceptible = graph_data$susceptible,
        infected = graph_data$infected,
        beta = beta,
        rho = 0.1,
        time = 50) %>%
    print_sir_values_2(paste0("Beta = ", beta)) %>% print()
}
```

Al disminuir $\beta$ ratio de transmisión vemos que los contagios prácticamente no se 'esparcen'. A medida que incrementamos
$\beta$ podemos ver que cada vez son las los que se contagian y más rápido.


En las observaciones anteriores hemos podido ver que para $\beta \epsilon {0.01, 0.1}$ los contagiados no superan a
la población susceptible.

Para $\beta = 0.3$ vemos que se contagia mayor parte de la población, hay un periodo donde
hay mas contagiados que susceptibles o recuperados pero, no llega a contagiar a toda la población. En algún punto la transmisión se corta.

Para $\beta = 0.5$ vemos que la población se contagia muy rápido y se llega a contagiar / recuperar toda.



## 2) Consider the random graph generated in the previous exercise.

### a) Plot its degrees distribution in linear and in log-log scale. Which is more helpful to understand this distribution?


```{r 2_a powerlaw_plot_function}
library(poweRlaw)
plot_degrees_distribution <- function(degree) {
  unique_degree <- unique(degree)

  m_pl <- displ$new(degree)

  d_est <- data.frame(K_min=sort(unique_degree)[1:(length(unique_degree)-2)], gamma=rep(0,length(unique_degree)-2), D=rep(0,length(unique_degree)-2))
  for (i in d_est$K_min){
    d_est[which(d_est$K_min == i),2] <- estimate_xmin(m_pl, xmins = i)$pars
    d_est[which(d_est$K_min == i),3] <- estimate_xmin(m_pl, xmins = i)$gof
  }
  K.min_D.min <- d_est[which.min(d_est$D), 1]

  #generate kmin & kmax pairs
  pairs <- as.data.frame(t(combn(sort(unique_degree), 2)))
  pairs$D <- rep(0, length(pairs$V1))
  pairs$gamma <- rep(0, length(pairs$V1))

  #scan D for all kmin-kmax pairs
  for (i in seq_along(pairs$D)){
    m_pl$setXmin(pairs[i,1])
    pairs[i,3]<- estimate_xmin(m_pl, xmin = pairs[i,1], xmax = pairs[i,2], distance = "ks")$gof
    pairs[i,4]<- estimate_xmin(m_pl, xmin = pairs[i,1], xmax = pairs[i,2], distance = "ks")$pars
  }

  #bs_pl_sat_cut <- bootstrap_p(m_pl, xmins = pairs[which.min(pairs$D), 1], xmax = pairs[which.min(pairs$D), 2], no_of_sims = 1000, threads = 8)
  k_sat <- pairs[which.min(pairs$D), 1]
  k_cut <- pairs[which.min(pairs$D), 2]

  #powerlaw
  m_pl <- displ$new(degree)
  est_pl <- estimate_xmin(m_pl, xmins = k_sat, xmax = k_cut, distance = "ks")
  m_pl$setXmin(est_pl)

  #lognormal
  m_ln <- dislnorm$new(degree)
  est_ln <- estimate_xmin(m_ln)
  m_ln$setXmin(est_ln)

  #exponential
  m_exp <- disexp$new(degree)
  est_exp <- estimate_xmin(m_exp)
  m_exp$setXmin(est_exp)

  #poisson
  m_poi <- dispois$new(degree)
  est_poi <- estimate_xmin(m_poi)
  m_poi$setXmin(est_poi)

  plot.new()
  plot(m_pl)
  lines(m_pl, col="red")
  lines(m_ln, col="green")
  lines(m_poi, col="blue")
  lines(m_exp, col="magenta")
  legend(x = "topright",
         fill = c("red", "green", "blue", "magenta"),
         legend = c("powerlaw", "lognormal", "exponential", "poisson"))
}

degree <- degree(graph_data$graph)
degree[degree>0] %>% plot_degrees_distribution()
```

Podemos comprobar que para nuestro grafo se adapta mejor la distribución *powerlaw* (linea roja).

## b) Does the degree distribution follows a Power Law? What if we consider only the nodes with degree above 5? (or 10? or 100?)
Como hemos visto en el apartado anterior la distribución de los grados sigue a *Power Law*.


Usaremos el contraste de hipótesis de *Kolmogorov-Smirnov* para definir si el subgrafo es o no libre de escala.
Diremos que $H_0 =$ libre de escala y $H_A = $ no libre de escala. Trataremos el coeficiente resultado como un p-valor
siendo $p-valor > 0.05$ no rechazaremos la hipótesis nula $H_0$, por lo que será libre de escala.

```{r 2_b1}
network.degree.distribution <- degree.distribution(graph_data$graph)
fit.scale.free <- power.law.fit(network.degree.distribution)
fit.scale.free[["KS.p"]]
```
`graph_data$graph` es un grafo libre de escala.

### con degree > 5

```{r 2_b2}
degree[degree>5] %>% plot_degrees_distribution()

V(graph_data$graph)$degree <- degree
graf2 <- subgraph(graph_data$graph,V(graph_data$graph)$degree > 5)
network.degree.distribution <- degree.distribution(graf2)
fit.scale.free <- power.law.fit(network.degree.distribution)
fit.scale.free[["KS.p"]]
```
`graph_data$graph` es un grafo libre de escala.

### con degree > 10

```{r 2_b3}
degree[degree>10] %>% plot_degrees_distribution()

graf2 <- subgraph(graph_data$graph,V(graph_data$graph)$degree > 10)
network.degree.distribution <- degree.distribution(graph_data$graph)
fit.scale.free <- power.law.fit(network.degree.distribution)
fit.scale.free[["KS.p"]]
```
`graph_data$graph` es un grafo libre de escala.

### con degree > 100

```{r 2_b4}
#degree[degree>100] %>% plot_degrees_distribution()
```
No hay nodos con grado superior a grado 100. No se puede ejecutar el algoritmo.


Vemos que a mayor grado las regresiones se "estiran".


## c) Find the best line that approximates the degree distribution after degree 10 (or 5?) using linear regression ( lm() ) on the log-log plane. Don't worry, it is almost all done in the following code. Explain in detail each line of the following code:

1. Crear distribución de grados de nuestro grafo
1. Quitar los nodos inconexos (grado == 0) y los primeros 10.
1. Montar log de los grados (Y) y de los id de cada nodo (X)
1. Montar una regresión lineal con el plano
1. Plot de la línea de regresión sobre XY

```{r 2_c}
degrees <- degree(graph_data$graph)
degrees_less_first10 <- which(degrees>0)[-(1:10)]  # remove the first 10 prob values
log_degrees_less_first10_y <- log(degrees[degrees_less_first10]) # degree
log_degrees_less_first10_x <- log(degrees_less_first10) # node id
LMI <- lm(log_degrees_less_first10_y~log_degrees_less_first10_x)$coefficients  # line coefficients
plot(degrees, pch=20, cex=0.7, xlab="Degree", ylab="Frequencies", main="degrees", log="xy")
points(exp(log_degrees_less_first10_x), exp(LMI[1]+LMI[2]*log_degrees_less_first10_x), col="red", type="l", lwd=2)
```


## d) What is the exponent of the Power Law for the degree probabilities?

Podemos comprobar que la correspondiente nube de puntos aparentemente se ajusta a una exponencial negativa.

```{r 2_d}
data.dist <- data.frame(k=0:max(degree), p_k=degree_distribution(graph_data$graph))
data.dist <- data.dist[data.dist$p_k>0,]
ggplot(data.dist) + geom_point(aes(x=k, y=p_k)) + theme_bw()
```

## 3) Use the routine sample_pa( ) to generate a rich-get-richer (preferential attachment) graph with similar degree distribution of  the *directed* facebook graph of the file **facebook_sample_anon.txt**. Use the code similar to: sample_pa(n.GRAPH, out.seq=degree(GRAPH,mode="out")) Plot the degree distribution of the generated graph (log-log). What is the exponent of the power law of the generated graph for the in-degrees?

Generate rich-get-richer
```{r 3_1}
links <- read.table("facebook_sample_anon.txt", header=FALSE, as.is=T)
facenet <- graph_from_data_frame(d=links, directed=FALSE)
graph <- sample_pa(n = length(V(facenet)), out.seq=degree(facenet, mode="out"))
```


```{r 3_2}
degrees <- degree(graph)
degrees[degrees>1000] %>% plot_degrees_distribution()
```


log-log
```{r 3_3}
degrees_less_first10 <- which(degrees>0)[-(1:10)]  # remove the first 10 prob values
log_degrees_less_first10_y <- log(degrees[degrees_less_first10]) # log degree
log_degrees_less_first10_x <- log(degrees_less_first10) # log node id
LMI <- lm(log_degrees_less_first10_y~log_degrees_less_first10_x)$coefficients  # line coefficients
plot(degrees, pch=20, cex=0.7, xlab="Degree", ylab="Frequencies", main="degrees", log="xy")
points(degrees_less_first10, exp(LMI[1]+LMI[2]*log_degrees_less_first10_x), col="red", type="l", lwd=2)
```


Podemos comprobar que la correspondiente nube de puntos aparentemente se ajusta a una exponencial negativa.

```{r 3_4}
data.dist <- data.frame(k=0:max(degrees), p_k=degree_distribution(graph))
data.dist <- data.dist[data.dist$p_k>0,]
ggplot(data.dist) + geom_point(aes(x=k, y=p_k)) + theme_bw()
```
